\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{parskip}
\setlength{\parindent}{0pt}
\usepackage{multicol}
\usepackage{enumitem}
\setlist[2]{nosep}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[dvipsnames]{xcolor}

\usepackage{tikz}
\tikzset{>=latex}

\title{ISC 4933/5935\\ 
Iterative and Direct Solvers for Linear Systems\\
Lecture Notes}
\author{Bryan Quaife and Eitan Lees}
\date{Spring, 2019}

\begin{document}

\maketitle

\section*{Course Information}

\begin{itemize}[label={}]
  \item Lectures:
    \begin{itemize}[label={}]
        \item MWF 10:10 - 11:00 AM
        \item Some Fridays we'll treat as a lab.
    \end{itemize}
    
  \item Instructor:
    \begin{itemize}[label={}]
        \item Bryan Quaife
        \item DSL 444
        \item bquaife@fsu.edu
        \item Office Hours - by appointment
    \end{itemize}
  
  \item TA:
    \begin{itemize}[label={}]
        \item Eitan Lees
        \item DSL 451C
        \item ejl15b@my.esu.edu \emph{or} elees@fsu.edu
        \item Office Hours - by appointment
    \end{itemize}
  
  \item Grade Breakdown:
    \begin{itemize}[label={}]
        \item Assignments: 50\%
        \item Final Project: 50\%
    \end{itemize}
    
  \item Prerequisites:
    \begin{itemize}[label={}]
        \item Linear Algebra
        \item Scientific Computing
    \end{itemize}
    
  \item Topics:
    \begin{itemize}[label={--}]
        \item Review of linear algebra
        \item Matrix-vector multiplication
        \item Iterative solvers
        \item Direct solvers
        \item Matrix Decomposition
            \begin{itemize}
                \item eigenvalue
                \item singular value decomposition
                \item LU - lower upper decomposition
                \item QR
            \end{itemize}
    \end{itemize}
    
\end{itemize}

\section{Introduction to Linear Systems}

A linear system is a set of equations of the form 

\begin{equation}
    \sum_{j=1}^n a_{ij} x_j  = b_i \qquad i=1, \ldots, m
\end{equation}

We are typically given $a_{ij}$ and one of $x_j$ ir $b_i$. Our job is to find the other. In matrix form, 
\begin{equation}
    A \Vec{x} = \Vec{b}
\end{equation}
where $A \in \mathbb{R}^{m \times n}$, $\Vec{x}\in \mathbb{R}^{n}$, $\Vec{b}\in \mathbb{R}^{m}$

\begin{equation}
    A = 
    \begin{bmatrix}
    a_{11} & a_{12} & \ldots & a_{1n}\\
    a_{21} & a_{22} & \ldots & a_{2n}\\
    \vdots &  & \ddots & \vdots\\
    a_{m1} & a_{m2} & \ldots & a_{mn}\\
    \end{bmatrix}
    ,\quad
    \Vec{x}  = \begin{bmatrix}x_1\\\vdots\\x_n\end{bmatrix}
    ,\quad
    \Vec{b}  = \begin{bmatrix}b_1\\\vdots\\b_m\end{bmatrix}
\end{equation}

Linear systems arise when we 
\begin{itemize}[label={--}]
    \item Discretize a linear differential equation or integral equation.
    \item Linearizing a nonlinear differential equation or a nonlinear integral equation.
    \item Interpolating data with basis functions such as polynomials, wavelets, Fourier expansions.
    \item Optimizing an objective function $f(\Vec{x})$. 
    \item Statistical operations such as lienar regression. 
    \item Solve an inverse problem. 
    \item Data sciences.
\end{itemize}

With $A\in\mathbb{R}^{m\times n}$, we have 3 cases
\begin{equation*}
    1)\, m < n \qquad 2)\, m > n \qquad 3)\ m=n
\end{equation*}

If $m<n$, we have fewer equations than unknowns. Such a system typically has infinitely many solutions, and such a system is said to be \emph{underdetermined}. It can also have no solutions. 

\underline{EX}:
\begin{align*}
    x_1 + x_2 + 2x_3 &= 0 & x_1 + x_2 - 2x_3 &= 0\\
    x_1 - 2x_2 + 3x_3 &= 1 & 2x_1 +2x_2 - 4x_3&=7\\
    \text{\textcolor{red}{infinitely many }}&\text{\textcolor{red}{solutions}}
    & \text{\textcolor{red}{no solutions}}
\end{align*}

Underdetermined problems arise, for example, in inverse problems. Here, there is typically some kind of field that is unknown, and we only have a few measurements. For example in seismic, the distribution of the formations underground are predicted from only a few available measurements. 

If we have $m>n$, we have more equations than unknowns. Such a system typically has no solution, and such a system is said to be \emph{overdetermined}. It can also have 1 solution or infinitely many solutions. 


\underline{EX}:
\begin{align*}
    x+y&=1 & x+y&=1 & x+y&=1 \\
    2x+2y&=2 & 2x+2y&=2 & x-y&=2 \\
    3x+3y&=3 & 3x+4y&=3 & 3x-y&=4 \\
    \text{\textcolor{red}{infinitely many }}&\text{\textcolor{red}{solutions}}
    & &\hspace{-3em}\text{\textcolor{red}{one solutions}}
    & &\hspace{-3em}\text{\textcolor{red}{no solutions}}
\end{align*}

Overdetermined problems arise, for example, in interpolation. Suppose we are given data points $(x_i, y_i),\, i=1, \ldots, m$ and we seek a linear function interpolating this data.  

Suppose the interpolant is $f(x) = c_0 + c_1 x$

\begin{align*}
    c_0 + c_1x_1 &= y_1 \\
    c_0 + c_1x_2 &= y_2 \\
    \vdots\\
    c_0+c_1x_m &= y_m
\end{align*}


\begin{center}
    \begin{tikzpicture}
\draw[->] (-.2,0) -- node[below]{x}(3,0);
\draw[->] (0, -.2) -- node[left]{y}(0, 3);

\draw [fill] (1.3112583208878787,1.0765211279204028) circle [radius=.05];
\draw [fill] (2.276383642890794,2.044674796484563) circle [radius=.05];
\draw [fill] (1.9163779073319989,1.6835130305468704) circle [radius=.05];
\draw [fill] (0.7212503291945787,0.8422314649775957) circle [radius=.05];
\draw [fill] (1.212503291945787,1.5422314649775957) circle [radius=.05];
\draw [fill] (2.638675593592325,2.357531828971839) circle [radius=.05];
\draw [fill] (1.923010531706664,1.416594971539452) circle [radius=.05];
\draw [fill] (2.2117959600493227,2.3689196263469596) circle [radius=.05];

\draw[dashed] (0,0) -- (3,2.8);

\end{tikzpicture}
\end{center}
Let's focus on the case $m=n$ and suppose that $A\Vec{x} = \Vec{b}$ has a unique solution for every $\Vec{b}$. The three main tasks to be done are 

\begin{enumerate}[{label=\arabic*)}]
    \item Compute $A\Vec{x}$ for some $\Vec{x}\in\mathbb{R}^n$
    \item Solve $A\Vec{x}=\Vec{b}$ for some $\Vec{b}\in\mathbb{R}^n$
    \item Decompose $A$ as 
    
    \begin{tabular}{ll}
         $A=LU$ & Lower-Upper Decomposition  \\
         $A=QR$ & where Q is orthogonal ($Q^T=Q^{-1}$), R is upper triangular\\
         $A=U\Sigma V^T$ & Singular Value Decomposition where $U, V$ are orthogonal, $\Sigma$ is diagonal\\
         $A=VDV^{-1}$ & Eigenvalue Decomposition\\
         $A=A_{(:, J)}X$ & Interpolative Decomposition
    \end{tabular}
\end{enumerate}

\end{document}
